---
title: "p8105_hw5_tl3196"
author: "Tianshu Liu"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	message = FALSE, 
  dpi = 300,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

Import and tidy data from csv files.

```{r import_tidy_data}
# generate file list
file_path = "./data/con_exp_data/"
file_list = list.files(path = file_path, full.names = FALSE)

# import and tidy data
con_exp_df =
  expand_grid(
    sample = file_list
  ) %>% 
  mutate(
    data = purrr::map(
      .x = file_list, 
      ~read_csv(file = str_c(file_path, .x))),
    sample = str_sub(sample, end = -5)
  ) %>% 
   separate(
     col = sample, sep = '_', 
     into = c("arm", "id"),
     remove = FALSE
  ) %>% 
  unnest(data) %>% 
  pivot_longer(
    cols = week_1:week_8,
    names_to = "week",
    names_prefix = "week_",
    values_to = "observation"
  )

con_exp_df
```

Make a spaghetti plot for the tidy data.

```{r spaghetti_plot}
# spaghetti plot
con_exp_df %>% 
  ggplot(aes(x = week, y = observation, group = sample, color = sample)) + 
  geom_point(aes(shape = arm), alpha = .7) + 
  geom_line(aes(linetype = arm), alpha = .7)
```

From the plot, we can identify that there are not significant difference between participants in the two arms at the beginning of the study. 
Participants in the experimental arm have significant increasing trend on observations, while participants in the control arm almost remain the same as the beginning. 
With this trend, the difference between the two arms becomes more and more significant. 
From the 7th week, all participants in the experimental arm have higher observation than all participants in the control arm.
Therefore, compared to the control arm, the experimental arm has significant increase in observations.

## Problem 2

Import data from `homicide-data.csv`.

```{r import_data}
homicide_df = read_csv("./data/homicide-data.csv")
homicide_df
```

The raw homicide dataset collected by the Washington Post covers ``r nrow(homicide_df)`` homicide cases in 50 large U.S. cities. Each homicide case is described by ``r ncol(homicide_df)`` variables, including ``r colnames(homicide_df)``. The variables are shown in the table below.

```{r data_description, echo = FALSE}
col_names = colnames(homicide_df)
col_types = c()
for (col in col_names){
  col_types = c(col_types, class(pull(homicide_df, col)))
}
description = c(
  "case id",
  "date of the homicide being reported",
  "last name of the victim",
  "first name of the victim",
  "race of the victim",
  "age of the victim",
  "gender of the victim",
  "city of the homicide",
  "state of the homicide",
  "latitude of the homicide's location",
  "longtitude of the homicide's location",
  "current arresting condition"
)

knitr::kable(
  tibble(col_names, col_types, description)
)
```


```{r city_state}
homicide_df = 
  homicide_df %>% mutate(city_state = str_c(city, ", ", state))

homicide_df
```

```{r summarize_by_city}
# summarize total number of homicide by cities
homicide_df %>% 
  group_by(city) %>% 
  summarise(
    n_homicide = n()
  )

# summarize number of unsolved homicide by cities
homicide_df %>% 
  filter(disposition %in% c("Closed without arrest", "Open/No arrest")) %>% 
  group_by(city) %>% 
  summarise(
    n_unsolved = n()
  ) 
```

Estimate the proportion unsolved homicides in Baltimore

```{r balt_prop_test}
balt_n_total = 
  homicide_df %>% 
  filter(city == "Baltimore") %>% 
  count() %>% 
  as.numeric()

balt_n_unsolved = 
  homicide_df %>% 
  filter(
    city == "Baltimore",
    disposition %in% c("Closed without arrest", "Open/No arrest")
  ) %>% 
  count() %>% 
  as.numeric()

balt_prop = prop.test(x = balt_n_unsolved, n = balt_n_total, conf.level = 0.95) 
balt_prop_df = 
  balt_prop %>% 
  broom::tidy()

estimate = pull(balt_prop_df, estimate)
conf_low = pull(balt_prop_df, conf.low)
conf_high = pull(balt_prop_df, conf.high)

tibble(
  city = "Baltimore",
  estimate, 
  ci = str_c("(", conf_low, ", ", conf_high, ")")
)
```

Iterate prop test for each city

```{r prop_test}
# iterate prop test for each city
prop_result_df = 
  homicide_df %>% 
  group_by(city) %>% 
  summarise(
    n_homicide = n(),
    n_unsolved = sum(disposition %in% c("Closed without arrest", "Open/No arrest"))
  ) %>% 
  mutate(
    test_result = purrr::map2(.x = n_unsolved, .y = n_homicide, 
                              ~prop.test(x = .x, n = .y, conf.level = 0.95) %>% broom::tidy())
  ) %>% 
  unnest(test_result) %>% 
  mutate(
    ci = str_c("(", conf.low, ", ", conf.high, ")")
  ) %>% 
  select(city, estimate,conf.low, conf.high, ci)

prop_result_df
```

```{r plot}
# plot estimated and CIs for each city
prop_result_df %>% 
  mutate(
    city = fct_reorder(city, estimate)
  ) %>% 
  ggplot(aes(x = city, y = estimate, color = city)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) + 
  theme(legend.position = "none") + 
  theme(axis.text.x = element_text(size=6, angle=45, hjust = 1)) + 
  labs(
    x = "City",
    y = "Estimated Proportion",
    title = "Estimated Proportions and CIs of Unsloved Homicides for Cities"
  )
```

